---
title: "Client project"
author: "[Xilai Fu]"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: style.css
    theme: united
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: false
    code_folding: hide
    code_download: true
bibliography: references.bib
link-citations: true
---
# Introduction
Our objective is to fit a logistic model based on model fitting(evaluated by AIC and Nagelkerke pseudo R2) and interpretability. Given a number of variables with missing values,it is of significance to predict **TB** versus **non-TB** in medical setting.
```{r Load libraries,warning=FALSE,message = FALSE}
library(Matrix)
library(Ckmeans.1d.dp)
library(RcppRoll)
library(dplyr)
library(ggplot2)
library(DT)
library(kableExtra)
library(Hmisc)
library(table1)
library(MASS)
library(DescTools)
library(gridExtra)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
```


```{r}
bio_dat <-read.csv("tb_data_aggregated.csv", na.strings=c("","NA"))
```

# Data preprocessing
Mandatory data manipulation:\
1: Change "INC" into NA\
2: Merge the levels Yes and yes for abscess_septation\
3: Merge the levels None with No for disk_inv\
4: Delete redundant variables **caseNo**,**dx_rad**,**dx_spine**\
5: Encode the levels rather than **TB** as **non-TB**\


Optional data manipulation:\
1: Introduce additional variables **new_t2_bone_intensity** and **new_disk_intensity** which combine the levels **Hyperintense** and **Mixed** as 1 and **Hypointense** and **Isointense** as 0\
2: Introduce additional variable **new_abscess_wall_thick** which encodes [>2mm] as 1 and [<=2mm] as 0\
3: Delete the levels **Sacral**,**CT** junction(only one case) for site_primary\
4: Combine the level **Paravertebral anterior**(only one sample) with **Paravertebral** for abscess_location_1
```{r warning=FALSE,message = FALSE}
## change "INC" into NA
bio_dat[bio_dat=="INC"] <- NA

## merge the levels Yes and yes
bio_dat[bio_dat$abscess_septation%in%c("yes"),]$abscess_septation <- "Yes"

## merge None with No
bio_dat[bio_dat$disk_inv%in%c("None"),]$disk_inv <-"No"

## delete redundent variables 
bio_dat[,c("caseNo","dx_rad","dx_spine")] <-NULL

## encode non-TB
bio_dat[bio_dat$dx_correct!="TB","dx_correct"]<-"non-TB"


## optional data manipulation
## site_primary: Sacral,CT junction(only one case)
## abscess_location_1: Paravertebral,Paravertebral anterior
bio_dat<-bio_dat[!(bio_dat$site_primary%in%c("Sacral","CT junction")),]
bio_dat[bio_dat$abscess_location_1 %in%c("Paravertebral","Paravertebral anterior"),]$abscess_location_1<-"Paravertebral"

##c("Hyperintense","Mixed"),] assigned to 1, c("Hypointense","Isointense") to 0, and "Destroyed" to NA
bio_dat$new_t2_disk_intensity<-bio_dat$t2_disk_intensity
bio_dat[bio_dat$new_t2_disk_intensity%in%c("Hyperintense","Mixed"),]$new_t2_disk_intensity <-"1"
bio_dat[bio_dat$new_t2_disk_intensity%in%c("Hypointense","Isointense"),]$new_t2_disk_intensity <-"0"
bio_dat[bio_dat$new_t2_disk_intensity %in% c("Destroyed"),]$new_t2_disk_intensity <-NA
bio_dat$new_t2_disk_intensity<-as.factor(bio_dat$new_t2_disk_intensity)

##c("Hyperintense","Mixed"),] assigned to 1, c("Hypointense","Isointense") to 0(t2_bone_intensity)
bio_dat$new_t2_bone_intensity<-bio_dat$t2_bone_intensity
bio_dat$new_t2_bone_intensity<-as.character(bio_dat$new_t2_bone_intensity)
bio_dat[bio_dat$new_t2_bone_intensity%in%c("Hyperintense","Mixed"),]$new_t2_bone_intensity <-"1"
bio_dat[bio_dat$new_t2_bone_intensity%in%c("Hypointense","Isointense"),]$new_t2_bone_intensity <-"0"
bio_dat$new_t2_bone_intensity<-as.factor(bio_dat$new_t2_bone_intensity)

## abscess wall thickness (abscess_wall_thick) by thresholding at 2mm: [>2mm]=1 and [<=2mm]=0
bio_dat$new_abscess_wall_thick<-bio_dat$abscess_wall_thick
bio_dat[bio_dat$new_abscess_wall_thick<=2&!(is.na(bio_dat$new_abscess_wall_thick)),]$new_abscess_wall_thick <-0
bio_dat[(bio_dat$new_abscess_wall_thick>2)&!(is.na(bio_dat$new_abscess_wall_thick)),]$new_abscess_wall_thick <-1
bio_dat$new_abscess_wall_thick<-as.factor(bio_dat$new_abscess_wall_thick)
```



# Exploratory data analysis

## Missing data visualization
```{r warning=FALSE,fig.align='center',message = FALSE}
## count of NAs column-wise
miss_summary <-apply(bio_dat,2,function(x) sum(is.na(x)))
na_summary<-as.data.frame(miss_summary)
miss_columns <- row.names(na_summary)
row.names(na_summary) <- seq(1,37,1)
na_summary_1<-cbind(miss_columns,na_summary)

## columns with missing values
na_summary_1<-cbind(na_summary_1,variable_type = sapply(bio_dat, class))
na_summary_1$variable_type<-sapply(na_summary_1$variable_type,function(x) ifelse(x %in% c("factor","character"),"factor","continuous"))

missing<-na_summary_1[miss_summary != 0, ]

## Visualization fo na values column-wise
ggplot(data = missing, aes(
  x = reorder(miss_columns, miss_summary),
  y = miss_summary,
  fill = variable_type
)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = miss_summary),
            vjust = 0, hjust = 0) + theme_light() + ylim(0, 50) +
  theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0,
      hjust = 0
    ),
    legend.position = c(0.88, 0.117)
  ) + labs(y = "Number of missing values", x = "Features") +
  ggtitle("Missing values column wise") +
  coord_flip()

row_missing<-as.data.frame(apply(bio_dat,1,function(x) round(sum(is.na(x))/length(x),digits = 2)))
colnames(row_missing)<-"row_missing_percentage"
row_missing<-cbind(row_number=rownames(row_missing),row_missing)
rownames(row_missing)<-NULL
kable(row_missing)%>%
add_header_above(c(" ", "Missing values percentage for each row" = 1))
```

## Feature selection 
1:Exclude the variables with a large proportion of missing values(**t2_cord_intensity**,**cord_signal_abnormal**, and **kyphotic_num**)\

2:Select the features with low p values\ 

3:abscess_soft_tissue_1 and abscess_septation are correlated with abscess_location_1 based on their chi-square p value so I decided to delete abscess_location_1 and abscess_septation\

4:After feature selection,we deleted the rows with missing values and we ended up with 12 variables and 82 rows.

**Tests for accessing the correlation between variables**\
1: Chi_square test for testing the correlation between categorical features and response\
2: Two sample t test between continuous features and response

### features and the reponse
```{r warning=FALSE,message = FALSE}
bio_dat$dx_correct_1<-factor(bio_dat$dx_correct, levels=c("TB", "non-TB", "P-value"), labels=c("TB", "non-TB", "P-value"))

rndr <- function(x, name, ...) {
    if (length(x) == 0) {
        y <- bio_dat[[name]]
        s <- rep("", length(render.default(x=y, name=name, ...)))
        if (is.numeric(y)) {
            p <- t.test(y ~ bio_dat$dx_correct_1)$p.value
        } else {
            p <- chisq.test(table(y, droplevels(bio_dat$dx_correct_1)))$p.value
        }
        s[2] <- sub("<", "&lt;", format.pval(p, digits=3, eps=0.001))
        s
    } else {
        render.default(x=x, name=name, ...)
    }
}

rndr.strat <- function(label, n, ...) {
    ifelse(n==0, label, render.strat.default(label, n, ...))
}


#str(subset(bio_dat,select = -c(dx_correct)))
## summary and correlation table for dataset
table1(~.|dx_correct_1, data=subset(bio_dat,select = -c(dx_correct)),render = rndr,type = "pvalues",droplevels = F,test=T)
```

### categorical variables
```{r warning=FALSE,message = FALSE}
options(scipen=999,digits=4)
bio_dat_update<-bio_dat[,c("dx_correct","post_inv_any","t2_disk_intensity","t2_bone_intensity","t2_endplate_intensity","disk_inv","abscess_soft_tissue_mass_defined","abscess_wall","kyphosis_deform","odema","loss_height","age","new_abscess_wall_thick","abscess_vol","new_t2_bone_intensity","abscess_wall_thick","abscess_septation")]

## convert character to factor
bio_dat_update[sapply(bio_dat_update,is.character)] <- lapply(bio_dat_update[sapply(bio_dat_update, is.character)],as.factor)

## Remove the rows with missing values
bio_dat_update_1<-bio_dat_update[complete.cases(bio_dat_update),]

cat_var<-bio_dat_update_1 %>% select_if(is.factor)
cat_names<-colnames(cat_var)

cat_corr<-as.data.frame(lapply(1:length(cat_names), function(i) mapply(function(x, y) chisq.test(table(cat_var[[x]],cat_var[[y]]))[["p.value"]],
       cat_names[i], 
       cat_names)),row.names
=cat_names,col.names = cat_names
)

kable(cat_corr, align = "lccrr",digits = 3, format.args = list(scientific = FALSE))
```

### continuous variables
```{r warning=FALSE,message = FALSE}
library(GGally)
con_var<-bio_dat_update_1 %>% select_if(is.numeric)
con_var %>% ggpairs()+theme(axis.text.x = element_text(vjust = 1, color = "black",font_size(12)))
```




# Logistic modelling
Our objective is to fit logistic models with original (**t2_bone_intensity** **abscess_wall_thick**) and modified c(**new_t2_bone_intensity** **new_abscess_wall_thick**) potentially with useful(significant) interaction terms. 

Models are sequentially selected by removing interaction terms one at a time(those with huge standard error are preferred).

Finally we arrived at two models based on the set of variables that includes c(**new_t2_bone_intensity** **new_abscess_wall_thick**).

## model1.1
```{r warning=FALSE,message = FALSE}
bio_dat_update1<-subset(bio_dat_update_1,select = -c(new_t2_bone_intensity,new_abscess_wall_thick)) # without new_t2_bone_intensity and new_abscess_wall_thick

bio_dat_update2<-subset(bio_dat_update_1,select = -c(t2_bone_intensity,abscess_wall_thick)) # without t2_bone_intensity and abscess_wall_thick

bio_dat_update_1<-bio_dat_update_1[bio_dat_update_1$disk_inv!="No",]
bio_dat_update1<-subset(bio_dat_update_1,select = -c(new_t2_bone_intensity,new_abscess_wall_thick)) # without new_t2_bone_intensity
bio_dat_update2<-subset(bio_dat_update_1,select = -c(t2_bone_intensity,abscess_wall_thick)) # without t2_bone_intensity


mod_1<-glm(formula = dx_correct ~ post_inv_any + disk_inv + kyphosis_deform + 
    odema + loss_height + age + new_abscess_wall_thick+ 
    new_t2_bone_intensity + abscess_septation + post_inv_any:kyphosis_deform, family = poisson(link = "log"), data = bio_dat_update2)

mod_2<-glm(formula = dx_correct ~ post_inv_any, family = poisson(link = "log"), data = bio_dat_update2)

#PseudoR2(mod_1,which ="Nagelkerke")
tab_model(mod_1,show.intercept = TRUE,
  show.est = TRUE,show.se = TRUE,show.aic=TRUE,collapse.ci = TRUE,show.r2 = FALSE,title ="Summary of the Best Model")


```

## model1.2
```{r warning=FALSE,message = FALSE}
mod_2<-glm(formula = dx_correct ~ post_inv_any + disk_inv + kyphosis_deform + 
    odema + loss_height + age + new_abscess_wall_thick + 
    new_t2_bone_intensity + abscess_septation + post_inv_any:kyphosis_deform + 
    post_inv_any:age + post_inv_any:new_t2_bone_intensity + 
    disk_inv:odema , family = binomial, data = bio_dat_update2)
summary(mod_2)
options(scipen=999,digits=4)
exp(cbind(OR =coef(mod_2),confint(mod_2)))



confint(mod_2)
coef(mod_2)
mod_2$coefficients
summary(mod_2)

z <-qnorm(0.975)
fitted_co <- coef(mod_2)[4]
se <- sqrt(diag(vcov(mod_2)))[4]
ci <- fitted_co + c(-z, z) * se
ci
summary(mod_2)
exp(fitted_co)

confint(mod_2)
exp(confint.default(mod_2))

a <-7.14
se <- 3.35
ci <- a + c(-z, z) * se
ci
exp(ci)

tab_model(mod_2)
#PseudoR2(mod_2,which ="Nagelkerke") 
```


# Conclusion {-}
Based on the output of the logistic models that we fitted, we preferred to go with/ 
model1.2 with an AIC = 41.29 and Nagelkerke pseudo R2 = 0.778 instead of model 1.1 with an AIC = 40.4 and pseudo R2 = 0.863 because model1.2 has more significant terms and terms with lower p-values in general than that of model1.2. in addition,model 1.1 has multiple insignificant interaction terms, which are problematic for inference.




